---
title: "Visualizing U.S. Midterm Turnout"
subtitle: "Over the Last 3 Midterms"
author: "Max McKitrick & Emily Dohler Rodas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
```

## Introduction

This section outlines the research idea. We can also cite related work here [@Wickham2014; @Baumer2017].

Note that compiled term paper (the PDF) is supposed to be more text-centered than the RMarkdown documents we used in class, i.e. the text sections are more detailed and big or redundant code chunks can be hidden.

## Data

This section describes the data sources and the data gathering process.

We found our data from several sources. We found the 2018 turnout data for each
state on "Statista.com," which did the work of gathering turnout data from each 
states Secretary of State office. Similarly, we found the 2014 midterm turnout
data from "electproject.org," which also collected the data from Secretary of
State offices. Finally,the 2022 turnout data is also from 
"electproject.orgthefulcrum.us." Finally, we found youth turnout data for 2014 &
2018 from "circle.tufts.edu," which were estimates collected from exit polls.

We then collected all of this data into one Excel spreadsheet. We organized the
spreadsheet such that the columns represented variables -- we had one with the 
state names, three with turnout for each midterm, and two for the youth turnouts
in 2014 and 2018. The code chunk below imports the Excel data we collected and
organized.
```{r}
# A code chunk that exemplifies the data gathering process
library(usmap)
library(ggplot2)
library(readxl)
library(readxl) 
SURV727_Final<-read_excel("C:/Users/maxmc/OneDrive/Desktop/SURV727_Final.xlsx")
View(SURV727_Final)
```

## Results

This section presents the main results.



### Analysis

This section presents the main results, such as (for example) stats and graphs that show relationships, model results and/or clustering, PCA, etc.

```{r}
## Visualizing the turnout in 2022
plot_usmap(data = SURV727_Final, values = "turnout_22", color = "black") + 
  scale_fill_continuous(name = "2022 Turnout Percentage", label = scales::comma) + 
  theme(legend.position = "right")
```

```{r}
# What happens here depends on the specific project
```

```{r}
# What happens here depends on the specific project
```

## Discussion

This section summarizes the results and may briefly outline advantages and 
limitations of the work presented.

## References
