---
title: "Visualizing U.S. Midterm Turnout"
subtitle: "Over the Last 3 Midterms"
author: "Max McKitrick & Emily Dohler Rodas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
```

## Introduction

Git Repo: https://github.com/MaxMcK412/TermPaper_727.git

Many political observers have pointed to an inflated youth turnout as the reason
for the Democrats’ strong performance in the 2022 midterm election results 
(Tufts). We do a top-line comparison of turnout in each midterm followed by a 
comparison of youth turnout between 2014 and 2018. We chose midterms because 
they generally draw less attention from the media and from word of mouth, so we
thought it would be more interesting. This did end up being a limiting factor 
due to a lack of data in some areas.

## Data

This section describes the data sources and the data gathering process.

We found our data from several sources. We found the 2018 turnout data for each
state on "Statista.com," which did the work of gathering turnout data from each 
states Secretary of State office. Similarly, we found the 2014 midterm turnout
data from "electproject.org," which also collected the data from Secretary of
State offices. Finally,the 2022 turnout data is also from 
"electproject.org" Finally, we found youth turnout data for 2014 &
2018 from "circle.tufts.edu," which were estimates collected from exit polls.

We then collected all of this data into one Excel spreadsheet. We organized the
spreadsheet such that the columns represented variables -- we had one with the 
state names, three with turnout for each midterm, and two for the youth turnouts
in 2014 and 2018. The code chunk below imports the Excel data we collected and
organized.
```{r}
# A code chunk that exemplifies the data gathering process
library(usmap)
library(ggplot2)
library(readxl)
library(readxl) 
SURV727_Final<-read_excel("C:/Users/maxmc/OneDrive/Desktop/SURV727_Final.xlsx")
View(SURV727_Final)
```

## Results

The topline results for turnout showed that 2014 had the lowest turnout, 36.4
percent, then 2022 had the next highest at 46.8 percent, and 2018 had the 
highest turnout at 49.4%. The youth turnout data shows us that compared to 2014,
2018 had exceptionally high youth voter turnout.


### Analysis

This section presents the main results, such as (for example) stats and graphs 
that show relationships, model results and/or clustering, PCA, etc.

```{r}
## Visualizing the turnout in 2022
plot_usmap(data = SURV727_Final, values = "turnout_22", color = "black") + 
scale_fill_continuous(low = "red", high = "green",
                      name = "2022 Turnout Percentage", label = scales::comma) + 
  theme(legend.position = "right")
```

```{r}
## Visualizing the turnout in 2018
plot_usmap(data = SURV727_Final, values = "turnout_18", color = "black") + 
scale_fill_continuous(low = "red", high = "green",
                      name = "2018 Turnout Percentage", label = scales::comma) + 
  theme(legend.position = "right")
```

```{r}
## Visualizing the turnout in 2014
plot_usmap(data = SURV727_Final, values = "turnout_14", color = "black") + 
scale_fill_continuous(low = "red", high = "green",
                      name = "2014 Turnout Percentage", label = scales::comma) + 
  theme(legend.position = "right")
```

```{r}
## Visualizing the difference between turnout in 2018 and 2014
SURV727_Final$diff_1418 <- SURV727_Final$turnout_18 - SURV727_Final$turnout_14

plot_usmap(data = SURV727_Final, values = "diff_1418", color = "black") + 
  scale_fill_continuous(low = "red", high = "green", 
                name = "14/18 Difference Percentage", label = scales::comma, 
                limits = c(-25, 25)) + 
  theme(legend.position = "right")
```

```{r}
## Visualizing the difference between turnout in 2018 and 2022
SURV727_Final$diff_1822 <- SURV727_Final$turnout_22 - SURV727_Final$turnout_18

plot_usmap(data = SURV727_Final, values = "diff_1822", color = "black") + 
  scale_fill_continuous(low = "red", high = "green", 
                  name = "18/22 Difference Percentage", label = scales::comma,
                  limits = c(-25,25)) + 
  theme(legend.position = "right")
```

```{r}
## Visualizing the difference between youth turnout in 2014 and 2018
SURV727_Final$youthdiff_1418 <- SURV727_Final$youth_turnout_14 - SURV727_Final$youth_turnout_18

plot_usmap(data = SURV727_Final, values = "youthdiff_1418", color = "black") + 
scale_fill_continuous(low = "red", high = "green", 
            name = "14/18 Youth Difference Percentage", label = scales::comma,
            limits = c(-25,25))+ 
  theme(legend.position = "right")
```

## Discussion

This section summarizes the results and may briefly outline advantages and 
limitations of the work presented.

There are a few notable things in the results. For instance, several states had
sharp decreases in turnout from 2018 to 2022 -- North Dakota, Alabama, Missouri,
Montana, and West Virginia. We see this because these states had competitive 
Senate races in 2018 but did not have competitive races in 2022; however, this
is not universally true. For example, despite having a competitive Senate Race
in 2018 and not in 2022, Indiana had higher voter turnout in 2022. Nearly all 
states had lower voter turnout in 2014 compared to 2018 but a couple exceptions
are Alaska and Colorado. Upon doing further research, you quickly find the 
reason for this is that Alaska had competitive Governor and Senate races in 2014
but not in 2018. Perhaps another point of analysis for studying voter turnout
could be the perception of competitiveness in races. Even though Colorado's
turnout in 2018 was lower than in 2014, the youth turnout exploded by nearly 75%
in 2018 over 2014. The strong preference for Democrats among youth voters can
help us explain the quick leftward trend of Colorado. The State with the most 
extreme increase in youth voter turnout was Montana -- about a 2.5 times 
increase from 2014 to 2018. On the other hand, Louisiana experienced nearly no
change in youth voter turnout from 2014 to 2018. 

Visualizing differences in voter turnout over midterms allows us to pinpoint 
which states are succeeding in getting voters to the polls, thus allowing us 
to isolate successful strategies. Visualizing youth voter turnout is important
for the same reason, but perhaps even more important since youth turnout is 
historically so abysmal.

Because the election was so recent, we did not have youth voter turnout 
available to us for the majority of States. There were several swing states that
had the results of exit polls published; however, there were not enough to give
us a full picture. In the few States we did see exit polls for, it seems like
youth turnout was probably slightly lower than 2018. 

## References
https://www.electproject.org/2022g
https://circle.tufts.edu/latest-research/2018-youth-voter-turnout-
increased-every-state
https://www.census.gov/topics/public-sector/voting.html